### 2024-04-12 new_workflow.Rmd
inital commit

---
title: "updated_workflow"
output: html_document
date: "2024-03-29"
---

```{r}
#library(mspms)
source("R/prepare_for_normalyzer.R")
source("R/extract_design_matrix.R")

### Loading the files ###
lfq_filename = "legacy/protein-peptides-lfq.csv"
#file "protein-peptides.csv" exported from PEAKS identification
id_filename = "legacy/protein-peptides-id.csv"

# Prepare the data for normalyzer analysis
prepared_data = prepare_for_normalyzer(lfq_filename,id_filename)

# Extracting design matrix
design_matrix = extract_design_matrix(prepared_data)

# normalyze the data

source("R/normalyze.R")

normalyzed_data = normalyze(prepared_data,design_matrix)

source("R/handle_outliers.R")
outliers = handle_outliers(normalyzed_data)

#outliers get converted to NAs. We dont want to impute these. We want to impute the values that are not NAs
source("R/impute.R")
imputed = impute(outliers)

# Writing sequences to file
peptide_library2 = peptide_library %>% 
  mutate(library_reference_id = gsub(">","",library_reference_id))

saveRDS(peptide_library2,"peptide_library")

source("R/join_with_library.R")

joined_with_library = join_with_library(imputed)

# Doing all the clevages

## Calculating where all of the cleavages are at.

source("R/nterm_cleavage.R")
source("R/cterm_cleavage.R")

source("R/add_cleavages.R")

final_data = add_cleavages(joined_with_library)

## Now that we have the final data, we need to

```