### 2024-04-12 mspms_shiny/about.Rmd
fixed a bunch of stuff.

---
title: "MSPMS-Shiny"
author: "Charlie Bayne"
date: "`r Sys.Date()`"
output:
  html_document:
        theme: cosmo
        code_folding: hide
        highlight: tango
        df_print: paged
---

```{r setup, include=FALSE}
library(tidyverse)
ggplot2::theme_set(ggprism::theme_prism(palette = "viridis"))
knitr::opts_chunk$set(echo = TRUE,warning = FALSE, message = FALSE)
```

# About

This is a Shiny app designed to use the [mspms R package](https://github.com/baynec2/mspms) to analyze data from the [Multiplex Substrate Profiling by Mass Spectrometry (MSP-MS) method](https://pubmed.ncbi.nlm.nih.gov/36948708/)

# How to use

On the side bar, you will see two places to upload files.

1. The first file should be the LFQ file from PEAKS. This file should be a csv file.
2. The second file should be the ID file from PEAKS. This file should also be a csv file.

You will also see two tabs at the top of the page, About and Output.. 
About links to these instructions.

Output contains a rendering of the normalized data, and a button that says download. This button will allow you to download the data that has been processed as a .csv file.

Be patient with the output tab, it takes a while for the code to execute. When it is finished, the table will appear and the data will be ready for download.