### 2024-04-16 mspms_shiny/about.Rmd
major changes

---
title: "MSPMS-Shiny"
author: "Charlie Bayne"
date: "`r Sys.Date()`"
output:
  html_document:
        theme: cosmo
        code_folding: hide
        highlight: tango
        df_print: paged
---

```{r setup, include=FALSE}
library(tidyverse)
ggplot2::theme_set(ggprism::theme_prism(palette = "viridis"))
knitr::opts_chunk$set(echo = TRUE,warning = FALSE, message = FALSE)
```

# About

This is a Shiny app designed to use the [mspms R package](https://github.com/baynec2/mspms) to analyze data from the [Multiplex Substrate Profiling by Mass Spectrometry (MSP-MS) method](https://pubmed.ncbi.nlm.nih.gov/36948708/)

# How to use

On the side bar, you will see three places to upload files.

1. The first file should be the LFQ file as exported from PEAKS. This file should be a csv file.

This file should look like the following:

```{r}
DT::datatable(head(mspms::peaks_lfq,2))
```

2. The second file should be the ID file as exported from PEAKS. This file should also be a csv file.

This file should look like the following:

```{r}
DT::datatable(head(mspms::peaks_id,2))
```

3. The third file should be the design matrix file. This should be a csv file.

The design matrix file should be a csv file with the following columns: "sample" and "group". The sample column should contain the names of the samples as found in the PEAKS files. The group column should contain the group that the sample belongs to. This can be any thing, as long as it is the same per group.

Here is an example of what the design matrix file should look like for the test files.

```{r}
DT::datatable(mspms::design_matrix)
```

You will also see two tabs at the top of the page, About and Output.. 
About links to these instructions.

Output contains a rendering of the normalized data, and a button that says download. This button will allow you to download the data that has been processed as a .csv file.

Be patient with the output tab, it takes a while for the code to execute. When it is finished, the table will appear and the data will be ready for download.