### 2024-04-16 troubleshooting.Rmd

---
title: "troubleshooting"
output: html_document
date: "2024-04-16"
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

```{r}

### Loading the files ###
lfq_filename = "test/protein-peptides-lfq_2.csv"
#file "protein-peptides.csv" exported from PEAKS identification
id_filename = "test/protein-peptides-id_2.csv"

# Prepare the data for normalyzer analysis
prepared_data = prepare_for_normalyzer(lfq_filename,id_filename)

design_matrix = readr::read_csv("test/design_matrix_2.csv")
normalyzed_data = normalyze(prepared_data,design_matrix)

outliers = handle_outliers(normalyzed_data,design_matrix)

imputed = impute(outliers)
joined_with_library = join_with_library(imputed)
final_data = add_cleavages(joined_with_library)

prepared_for_stats = prepare_for_stats(final_data,design_matrix)

anova = mspms_anova(prepared_for_stats)

ttests = mspms_t_tests(prepared_for_stats)

```