### 2024-05-31 inst/rmarkdown/templates/mspms_report/skeleton/skeleton.Rmd
major changes

---
title: "MSPMS Report"
date: "`r Sys.Date()`"
output:
  html_document:
        theme: cosmo
        toc: true
        toc_depth: 6
        toc_float:
            collapsed: true
            smooth_scroll: true
        code_folding: hide
        highlight: tango
        df_print: paged
params:
  prepared_data: NULL
  design_matrix: NULL
  peptide_library: mspms::peptide_library
  n_residues: 4
---

```{r setup, include=FALSE}
library(dplyr)
ggplot2::theme_set(ggplot2::theme_bw())
knitr::opts_chunk$set(echo = TRUE,warning = FALSE, message = FALSE)
```

# Info

This report has been automatically generated by the mspms package. It contains standard results of the mspms analysis pipeline.

R session info is as follows:

```{r}
sessionInfo()
```
# Data Processing

```{r}
# The standard mspms data processing pipeline is used here. 
mspms_data = mspms::mspms(prepared_data,design_matrix)
```

# Overview 
```{r}
mspms::plot_pca(mspms_data)
```

**Figure 1.** Here a PCA of the experiment is displayed. The color represents time, while shape and linetype of eclipse (if there are enough points to calculate) represents condition.

```{r,fig.height = 8, fig.width = 10}
mspms::plot_heatmap(mspms_data)
```

**Figure 2.** Here an interactive heatmap of the experiment is displayed. Mouse over to see more information on each cell. Column side colors represent condition and time.

# Statistics

```{r}
t1 = mspms::mspms_anova(mspms_data) %>% 
  dplyr::arrange(p.adj)

DT::datatable(t1)
```

**Table 1. ** Here anova results are displayed for each condition (testing to see if there is an effect on time).

```{r}
t2 = log2fc_t_test(mspms_data) %>% 
  dplyr::arrange(p.adj)

DT::datatable(t2)
```

**Table 2.** Here T-test results are displayed as compared to time 0 for each condition

```{r}
plot_volcano(t2)
```

**Figure 3.** Here the Log2fc and T tests are displayed as volcano plots relative to T0 for each condition.

# Cleavage Motifs

```{r,fig.height = 10, fig.width = 6}
print(mspms::plot_all_icelogos(mspms_data))
```

**Figure 4.** Here the cleavage motifs enriched in each condition are shown.