### 2024-06-04 README.Rmd

---
output:
  github_document: default
  self_contained: default
---

<!-- README.md is generated from README.Rmd. Please edit that file -->

```{r, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>",
  cache = FALSE
)

ggplot2::theme_set(
  ggplot2::theme_minimal()
)

```

# mspms

<!-- badges: start -->
[![R-CMD-check](https://github.com/baynec2/mspms/actions/workflows/R-CMD-check.yaml/badge.svg)](https://github.com/baynec2/mspms/actions/workflows/R-CMD-check.yaml)
[![Codecov test coverage](https://codecov.io/gh/baynec2/mspms/branch/main/graph/badge.svg)](https://app.codecov.io/gh/baynec2/mspms?branch=main)
[![BioC status](http://www.bioconductor.org/shields/build/release/bioc/mspms.svg)](https://bioconductor.org/checkResults/release/bioc-LATEST/mspms)
<!-- badges: end -->

The goal of mspms is provide a concise code-base for the normalization and 
data processing required to analyze data from the 
[Multiplex Substrate Profiling by Mass Spectrometry (MSP-MS) method](https://pubmed.ncbi.nlm.nih.gov/36948708/).

Additionally, we provide a [graphical user interface powered by shiny apps](https://gonzalezlab.shinyapps.io/mspms_shiny/) that allows for a user to
utilize the method without requiring any R coding knowledge.

## Installation

You can install the development version of mspms from github.

```{r,eval = FALSE}
devtools::install_github("baynec2/mspms",
                         ref = "develop")
```

## Quickstart

To generate a general report using your own data, run the following code. 
It requires data that has been prepared for mspms data analysis by a 
converter function, and a design matrix. For more information see subsequent 
sections.

```{r,eval = FALSE}
mspms::generate_report(
  prepared_data = mspms::peaks_prepared_data,
  design_matrix = mspms::design_matrix,
  outdir = "../Desktop/mspms_report"
)
```

The above command will generate a .html file containing a general mspms analysis.

There is much more that can be done using the mspms package- see the following 
sections for more information.

## Overview

There are 4 different types of functions in this package.

1. Making mspms generically useful. These functions are focused on making mspms 
more generally useful to a wider audience outside of the very specific workflow
traditionally used in the O’Donoghue lab. Allows for the use of other types of
upstream proteomic data processing, different peptide libraries, etc.

2. Data processing/ normalization. These functions allow the user to normalize
and process the MSP-MS data.

3. Statistics. These methods allow the user to perform basic statistics on the 
normalized/processed data.

4. Data visualization. These functions allow the user to visualize the data.

**Making mspms generically useful**.   
1. *prepare_peaks()*:Takes two input files from PEAKS and combines them. 
2. *prepare_pd()*: prepares exported files from proteome discoverer.  
3. *calculate_all_cleavages()*: Calculates all possible cleavages for peptide
library sequences.

**Data Processing/ Normalization**.  
1. *normalyze()*: Normalizes values.     
2. *handle_outliers()*: Looks for outliers across replicates. Removes them.       
3. *impute()*: Imputes data for missing values (not including NAs introduced 
by handle_outliers()).  
4. *join_with_library()*: Joins the normalized data with the peptide library 
sequences.  
5. *add_cleavages()*: Figures out the locations of the detected cleavages within
the library of peptide sequences used. Is it cleaved at the N or C terminus, 
or both? Uses *cterm_cleavage()* and *nterm_cleavage()* to do this.      
6. *polish()*: combines the cleavage info into one column. Discards peptides
that were cleaved on both sides or not at all.  
7. *prepare_for_stats()*: Reshapes the data into a long format and appends the 
data in the design matrix to make statistical testing easy to perform.  
8 *mspms():* combines all of the above functions into one convenient function.

**Calculations/ Statistics**.   
1. *mspms_log2fc()*: Calculates the log2 fold change across experimental 
conditions relative to T0.  
2.*mspms_t_tests()*: Conducts T-tests to determine if there are any peptides 
that are significantly different across experimental conditions relative to time 
0.  
3. *log2fc_t_test()*: Combines results from mspms_log2fc() and mspms_t_tests() 
into one convenient tibble.   
4. *log2fct_condition()*: Calculates the log2 fold change and t tests for the 
specified condition at a user specified time.    
5.*log2fct_time()*: Calculates the log2 fold change and t tests between the 
specified times within a user specified time.   
6. *mspms_anova()*: Conducts ANOVA tests to determine if there are any peptides 
that are significantly different across all time points.   
7. *count_cleavages_per_pos():* Counts the number of cleavages at each position 
of the library. Useful for determining endoprotease vs exoprotease activity.

**Data Visualization**.   
1. *plot_heatmap()*: Conducts hierarchical clustering analysis and plots an
interactive heatmap to visualize overall patterns in the data.    
2. *plot_pca()*: PCA analysis to visualize the data in a 2D space.  
3. *plot_time_course()*: Plot peptides over time by condition.  
4. *plot_cleavages_per_pos()*: Plot the number of cleavages at each position 
in the library.Useful for determining endoprotease vs exoprotease activity.

**Icelogo**.

1. *calc_AA_count_of_motif()*: calculates the count of each amino acid at each
position in a vector of motifs.  
2. *calc_AA_prop_of_motif()*: calculates the proportion of each amino acid at 
each position from the output of calc_AA_count_of_motif().  
3. *calc_AA_motif_zscore()*: calculates z scores for each amino acid at each 
position in a vector of motifs.  
4. *calc_AA_percent_difference()*: calculates the percentage difference of 
5. *prepare_sig_p_dif()*: prepares the data for plotting percent difference 
icelogos. Removes AAs at positions that are not significantly different.  
6. *prepare_fc()*: prepares the data for plotting fold change icelogos.
Removes AAs at positions that are not significantly different.  
7. *plot_pd_icelogo()*: plots a percent difference icelogo.  
8. *plot_fc_icelogo()*: plots a fold change ice logo.  
9. *plot_icelogo()*:  wraps plot_pd_icelogo() and plot_fc_icelogo() into 
one function.  
10. *plot_all_icelogos()*: convenience function to plot all icelogos relative to 
time 0 from an experiment.    
**Reports**. 
1. *generate_report()*: Generates a simple analysis of the mspms data that 
should be generically applicable to all mspms experiments.

## Making Generically Usefull.

### Formating Peaks File Outputs.

All analysis performed by this package is downstream of the data generated by 
the PEAKS software. The first step is to combine the two files generated by
PEAKS.

The files coming from peaks should be generated according to the instructions 
found [here](www/PeaksDataAnalysis_howto.pdf).

```{r}
library(dplyr)
library(mspms)

### Loading the files ###
lfq_filename <- "tests/testdata/protein-peptides-lfq.csv"
# file "protein-peptides.csv" exported from PEAKS identification
id_filename <- "tests/testdata/protein-peptides-id.csv"

# Prepare the data for normalyzer analysis
peaks_prepared_data <- prepare_peaks(lfq_filename, id_filename)
```

### Formating Proteome Discoverer File Outputs.

We can prepare proteome discover files for analysis with mspms as follows.

```{r}
prepared_proteome_discoverer <- prepare_pd("tests/testdata/proteome_discoverer_output.xlsx")
```

### Calculating all cleavages

We might want to calculate all possible cleavages for the peptide library 
sequences. This is useful for downstream analysis, especially when we are 
looking at the specificity of the cleavage sites via plot_cleavage_motif() as
this requires a background of all possible cleavages

We can do this by specifying the number of amino acids after the cleavage site 
that we are interested in. First lets try 4, which is the default.

```{r}
all_peptide_sequences <- mspms::calculate_all_cleavages(mspms::peptide_library$library_real_sequence,
  n_AA_after_cleavage = 4
)
head(all_peptide_sequences)
```

We could also try 5 AA after each cleavage site

```{r}
all_peptide_sequences <- mspms::calculate_all_cleavages(mspms::peptide_library$library_real_sequence,
  n_AA_after_cleavage = 5
)
head(all_peptide_sequences)
```

## Data Normalization/ Processing

### Loading design matrix

Before we normalyze data, we need to know what samples are in what groups. We can do that by defining a design matrix. 
```{r}
design_matrix <- readr::read_csv("tests/testdata/design_matrix.csv")

head(design_matrix)
```
### Normalyzing data

mspms uses the NormalyzerDE package to do normalization under the hood.

We can normalyze the data as follows:  
```{r}
normalyzed_data <- normalyze(peaks_prepared_data, design_matrix)

```

### Handling Outliers

Here we use a Dixon test from the outliers package to detect outliers from each 
of our replicates. We need to know what samples are part of which groups, so we 
need to specify the design matrix here too. Make sure that the column header 
names are "sample" and "group" just like before.

```{r}
design_matrix <- readr::read_csv("tests/testdata/design_matrix.csv")
outliers <- handle_outliers(normalyzed_data, design_matrix)

```

### Imputation of data

We have a lot of missing, or 0 values. For these, we need to impute them so we 
can do downstream statistics.

```{r}
imputed <- mspms::impute(outliers)

```

### Joining with Library

Next we need to join everything with the sequences of the peptide library.  
```{r}
joined_with_library <- mspms::join_with_library(imputed)
```

### Calcuating cleavages.

Next, we need to determine where the peptide sequences are cleaved.

We check both the N and C terminus.

Sequences are presented as the user specified number of amino acids on 
both sides of a cleavage. The default is 4, but there is interest in looking at
motifs that are farther away from the cut site. Note that X indicates that there
was nothing at that position because it is over the end.

```{r}
cleavage_added_data <- mspms::add_cleavages(joined_with_library, n_residues = 4)

head(cleavage_added_data)
```

### Adding metadata back

We need to add information about our samples back to perform downstream 
statistics on the data. We accomplish this with the prepare for stats file.

```{r}
prepared_for_stats <- mspms::prepare_for_stats(cleavage_added_data,
                                               design_matrix)
```

### Polishing

This function polishes the data by combining the cterm and nterm cleavage 
information into one column while removing any rows that don't have any 
cleavage information or have cleavage information on the cterm and nterm.

```{r}
polished_data <- mspms::polish(cleavage_added_data)
```

### Complete workflow

There are a lot of steps to get to the final normalized data we listed above. 
We have combined all of these steps into one function, mspms() for convenience.
This offers a nice balance between allowing the user the option to use 
individual steps and providing a quick way to get to the final data.

```{r}
mspms_data <- mspms::mspms(peaks_prepared_data, design_matrix)
```

## Statistics

mspms provides a number of convenience functions to conduct statistics
on the data.

First, we need to prepare the data for stats. This involves reshaping the data 
into the long format so it is easy to conduct statistics on and then appending 
the data in the design matrix that allows us to determine what sample contains 
what conditions/time and conduct the appropriate statistics.

Now we can conduct the statistics:

### T tests

T tests are performed within each condition and compared to time 0.

For example, for an experiment where there are two conditions, 
DMSO and MZB as well as 3 time points, 0, 1, and 2, the t tests would be
as follows:

* DMSO.TO vs DMSO.T0. 
* DMSO.T0 vs DMSO.T1. 
* DMSO.T0 vs DMSO.T2.

* MZB.T0 vs MZB.T0. 
* MZB.T0 vs MZB.T1. 
* MZB.T0 vs MZB.T2.

```{r}
# Perform T test
t_test_stats <- mspms::mspms_t_tests(mspms_data)
```

### log2FC

We can also calculate the log 2 fold change. 
The comparisons here are the same as for the T tests.  
```{r}
# calculate log2fc
log2fc <- mspms::mspms_log2fc(mspms_data)

head(log2fc)
```

### log2fc_t_tests.

Oftentimes (such as when you want to make volcano plots) it is most useful to
look at the log2fc and the t test statistics at the same time.

```{r}
log2fc_t_test <- mspms::log2fc_t_test(mspms_data)

head(log2fc_t_test)
```

### log2fct_condition

You may wish to compare two conditions at a specific time point. 
This function allows this

```{r}
condition_t = mspms::log2fct_condition(mspms_data,
                                       ref_condition = "DMSO",
                                       comparison_condition = "MZB",
                                       at_time = 60)

head(condition_t)

```

### log2fct_time

You may wish to compare two time points within a specific condition. 
This function allows

```{r}
time_t = mspms::log2fct_time(mspms_data,
                             within_condition = "DMSO",
                             ref_time = 0,
                             comparison_time = 60)

head(time_t)
```

### ANOVA

We also might want to perform an anova. Here, we have it set up to show the 
effect of time within each condition.

For example, for an experiment where there are two conditions, DMSO and MZB as
well as 3 time points, 0, 1, and 2, the anova would statistics for the effect of
time for each peptide within DMSO or MZB.

```{r}
# Doing ANOVA
anova_stats <- mspms::mspms_anova(mspms_data)

head(anova_stats)
```

## Common Data Visualizations

We also provide some functions that make common data visualizations easier.

### volcano plots

```{r}
volcano_plot <- mspms::plot_volcano(log2fc_t_test)

volcano_plot
```

### Plotting sequence specificity motif

Here use an approach similar to what is implemented in ICELOGO to visualize the
sequence specificity of the cleavage sites. This was used as reference to build 
the code: https://iomics.ugent.be/icelogoserver/resources/manual.pdf.

Below we can see this applied to see what cleavage sequences are enriched among
the peptides significantly different between time 0 and time 60 
within DMSO.

```{r}
cleavage_seqs <- time_t %>%
  dplyr::filter(log2fc > 3, p.adj <= 0.05) %>%
  dplyr::pull(cleavage_seq)

background_universe <- mspms::all_possible_8mers_from_228_library

mspms::plot_icelogo(cleavage_seqs, background_universe)
```

We could also look at the fold change instead of the percent difference

```{r}
mspms::plot_icelogo(cleavage_seqs, background_universe, type = "fold change")
```

We also provide a convenience function for looking at all icelogos relative to 
time 0 within the experiment.

This function includes cleavages that were significant at any time point once in 
the experimental set of the ICELOGO.

```{r}
mspms::plot_all_icelogos(mspms_data)
```

### PCA.

We can generate a PCA plot to visualize the data. 
Here, the colors show the different time points while the shape shows 
the different conditions.

```{r}
mspms::plot_pca(mspms_data)
```

### Hierchical clustering

We can also generate an interactive heatmap with the hierarchical clustering 
results.

The format of this readme does not allow for interactivity, so a static picture 
of the output is shown instead.

```{r,eval = FALSE}
mspms::plot_heatmap(mspms_data)
```

![](plot_heatmap_output.png)

### Ploting time course.

We also provide a function for plotting the mean intensity and standard 
deviation over time for each peptide in the data set by conditions.
Facets show peptide, color shows the condition.

This can be used in combination with the ANOVA function.
Below, we will use anova to calculate the peptides where we see an 
effect of time (in the DMSO group), and then plot these.

```{r}

top10sig <- anova_stats %>%
  dplyr::arrange(p.adj) %>%
  dplyr::pull(Peptide) %>%
  head(10)

p1 <- mspms_data %>%
  dplyr::filter(Peptide %in% top10sig) %>%
  mspms::plot_time_course()

p1
```

## Calculating the number of cleavages at each position in the library

We can also calculate the number of cleavages at each position of the library.
Here we will do this for peptides significantly different relative to time 0 for
DMSO and MZB.

```{r}
# Performing T tests
t_tests = log2fc_t_test(mspms_data) %>% 
  dplyr::filter(p.adj < 0.05,
                log2fc > 3)

count = mspms::count_cleavages_per_pos(t_tests)

p1 = mspms:::plot_cleavages_per_pos(count)

p1
```