### 2024-06-17 README.Rmd
updating README to include new functions.

---
output: github_document

---

<!-- README.md is generated from README.Rmd. Please edit that file -->

```{r, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>"
)

ggplot2::theme_set(
  ggplot2::theme_minimal()
)

```

# mspms

<!-- badges: start -->
[![R-CMD-check](https://github.com/baynec2/mspms/actions/workflows/R-CMD-check.yaml/badge.svg)](https://github.com/baynec2/mspms/actions/workflows/R-CMD-check.yaml)
[![Codecov test coverage](https://codecov.io/gh/baynec2/mspms/branch/main/graph/badge.svg)](https://app.codecov.io/gh/baynec2/mspms?branch=main)
[![BioC status](http://www.bioconductor.org/shields/build/release/bioc/mspms.svg)](https://bioconductor.org/checkResults/release/bioc-LATEST/mspms)
<!-- badges: end -->

The goal of mspms is provide a concise code-base for the normalization and 
data processing required to analyze data from the 
[Multiplex Substrate Profiling by Mass Spectrometry (MSP-MS) method](https://pubmed.ncbi.nlm.nih.gov/36948708/).

Additionally, we provide a 
[graphical user interface powered by shiny apps](https://gonzalezlab.shinyapps.io/mspms_shiny/) 
that allows for a user to utilize the method without requiring any R coding
knowledge.

## Installation

You can install the development version of mspms from github.

```{r,eval = FALSE}
devtools::install_github("baynec2/mspms",
                         ref = "develop")
```

## Quickstart

To generate a general report using your own data, run the following code. 
It requires data that has been prepared for mspms data analysis by a 
converter function, and a design matrix. For more information see subsequent 
sections.

```{r,eval = FALSE}
mspms::generate_report(
  prepared_data = mspms::peaks_prepared_data,
  design_matrix = mspms::design_matrix,
  outdir = "../Desktop/mspms_report"
)
```

The above command will generate a .html file containing a general mspms analysis.

There is much more that can be done using the mspms package- see the following 
sections for more information.

## Overview

There are 4 different types of functions in this package.

1. Pre-processing data. These functions are focused on making mspms 
more generally useful to a wider audience outside of the very specific workflow
traditionally used in the O’Donoghue lab. Allows for the use of other types of
upstream proteomic data processing, different peptide libraries, etc.

2. Data processing/ normalization. These functions allow the user to normalize
and process the MSP-MS data.

3. Statistics. These methods allow the user to perform basic statistics on the 
normalized/processed data.

4. Data visualization. These functions allow the user to visualize the data.

**Preprocessing data**.   
1. *prepare_peaks()*:Takes two exported files from PEAKS software, 
combines them, and prepares for processing by the mspms R package. 
2. *prepare_pd()*: prepares exported files from proteome discoverer.  
3. *calculate_all_cleavages()*: Calculates all possible cleavages for peptide
library sequences.

**Data Processing/ Normalization**.  
1. *normalyze()*: Normalizes values.     
2. *handle_outliers()*: Looks for outliers across replicates. Removes them.       
3. *impute()*: Imputes data for missing values (not including NAs introduced 
by handle_outliers()).  
4. *join_with_library()*: Joins the normalized data with the peptide library 
sequences.  
5. *add_cleavages()*: Figures out the locations of the detected cleavages within
the library of peptide sequences used. Is it cleaved at the N or C terminus, 
or both? Uses *cterm_cleavage()* and *nterm_cleavage()* to do this.      
6. *polish()*: combines the cleavage info into one column. Discards peptides
that were cleaved on both sides or not at all.  
7. *prepare_for_stats()*: Reshapes the data into a long format and appends the 
data in the design matrix to make statistical testing easy to perform.  
8 *mspms():* combines all of the above functions into one convenient function.

**Calculations/ Statistics**.   
1. *mspms_log2fc()*: Calculates the log2 fold change across experimental 
conditions relative to T0.  
2.*mspms_t_tests()*: Conducts T-tests to determine if there are any peptides 
that are significantly different across experimental conditions relative to time 
0.  
3. *log2fc_t_test()*: Combines results from mspms_log2fc() and mspms_t_tests() 
into one convenient tibble.   
4. *log2fct_condition()*: Calculates the log2 fold change and t tests for the 
specified condition at a user specified time.    
5.*log2fct_time()*: Calculates the log2 fold change and t tests between the 
specified times within a user specified time.   
6. *mspms_anova()*: Conducts ANOVA tests to determine if there are any peptides 
that are significantly different across all time points.   
7. *count_cleavages_per_pos():* Counts the number of cleavages at each position 
of the library. Useful for determining endoprotease vs exoprotease activity.
8. *count_cleavages_per_pos()*: Counts the number of cleavages at the position
of the initial peptide detected peptides were cleaved from.

**Icelogo calculations**

1. *calc_AA_count_of_motif()*: calculates the count of each amino acid at each
position in a vector of motifs.  
2. *calc_AA_prop_of_motif()*: calculates the proportion of each amino acid at 
each position from the output of calc_AA_count_of_motif().  
3. *calc_AA_motif_zscore()*: calculates z scores for each amino acid at each 
position in a vector of motifs.  
4. *calc_sig_zscore()*: given a matrix of zscores, determines which are significant given a user defined p value. 
5. *calc_AA_percent_difference()*: calculates the percentage difference of 
6. *calc_AA_fold_change()*: calculates the fold change of a matrix containing 
the proportions of amino acids at each position relative to a background matrix. 
7. *prepare_sig_p_dif()*: prepares the data for plotting percent difference 
icelogos. Removes AAs at positions that are not significantly different.  
8. *prepare_pd()*:prepares a matrix with significant amino acid percent 
differences by position for plotting.
9. *prepare_fc()*: prepares the data for plotting fold change icelogos.
10. *prepare_icelogo_data()*: uses all the functions described above to 
prepare the data for plotting icelogos. 
11. *extract_re()*: extracts regular expressions specifying amino acids above a
defined threshold cutoff for each position of a cleavage motif from an icelogo 
matrix.

**Data Visualization**.   
1. *plot_heatmap()*: Conducts hierarchical clustering analysis and plots an
interactive heatmap to visualize overall patterns in the data.    
2. *plot_pca()*: PCA analysis to visualize the data in a 2D space.  
3. *plot_time_course()*: Plot peptides over time by condition.  
4. *plot_cleavages_per_pos()*: Plot the number of cleavages at each position 
in the library.Useful for determining endoprotease vs exoprotease activity.
5.*plot_icelogo()*: Plot an icelogo given an icelogo matrix. 
6.*plot_all_icelogos()*: generates icelogos representing enriched sequences 
relative to time 0 for each conditions. This function compares cleavage 
sequences corresponding to peptides that are significant 
(p.adj < 0.05 and log2 fold change > 3) to a background of all possible
sequences present in the initial peptide library.

**Reports**. 
1. *generate_report()*: Generates a simple analysis of the mspms data that 
should be generically applicable to all mspms experiments.

## Data Pre-processing

### Formating Peaks File Outputs.

Analysis downstream of data exported from PEAKS software is supported.

exported files are generated in the O'donoghue lab as follows:

1. Create a New Project in PEAKS.  
2. Select data to add to PEAKS project.  
3. Add data, rename samples, set correct parameters. 
  • Highlight all the added data.  
  • Select Create new sample for each file. 
  • Select appropriate instrument and fragmentation method. 
  •  Set Enzyme to “None”. 
  • Rename samples according to enzymes, time points, and replicates. 
      • Can also ensure appropriate data is selected for these   
  • Select “Data Refinement” to continue. 
4. Data Refinement. 
  • On the top right, select MSP-MS as the predefined parameters. 
  •  Parameters shown are what should be used for MSP-MS data analysis. 
  •  Select “Identification” to continue. 
5. Peptide Identification. 
  • On the top right, select MSP-MS as the predefined parameters. 
  • Parameters shown are what should be used for MSP-MS data analysis. 
  • Identification -> Select Database -> TPD_237. 
  • Identification -> no PTMs. 
  • Can remove the PTMs by highlighting them and selecting Remove  
6. Label Free Quantification. 
  • Make sure Label Free is selected. 
  • Group samples -> add quadruplicates of samples to new group.

```{r}
library(dplyr)
library(mspms)

### Loading the files ###
lfq_filename <- "tests/testdata/protein-peptides-lfq.csv"
# file "protein-peptides.csv" exported from PEAKS identification
id_filename <- "tests/testdata/protein-peptides-id.csv"

# Prepare the data for normalyzer analysis
peaks_prepared_data <- prepare_peaks(lfq_filename, id_filename)
```

### Formating Proteome Discoverer File Outputs.

We can prepare proteome discover files for analysis with mspms as follows.

```{r}
prepared_proteome_discoverer <-
  prepare_pd("tests/testdata/proteome_discoverer_output.xlsx")
```

### Calculating all cleavages

We might want to calculate all possible cleavages for the peptide library 
sequences. This is useful for downstream analysis, especially when we are 
looking at the specificity of the cleavage sites via plot_cleavage_motif() as
this requires a background of all possible cleavages

We can do this by specifying the number of amino acids after the cleavage site 
that we are interested in. First lets try 4, which is the default.

```{r}
all_peptide_sequences <- mspms::calculate_all_cleavages(
  mspms::peptide_library$library_real_sequence,
  n_AA_after_cleavage = 4
  )
head(all_peptide_sequences)
```

We could also try 5 AA after each cleavage site

```{r}
all_peptide_sequences <- mspms::calculate_all_cleavages(
  mspms::peptide_library$library_real_sequence,
  n_AA_after_cleavage = 5
)
head(all_peptide_sequences)
```

## Data Normalization/ Processing

### Loading design matrix

Before we normalyze data, we need to know what samples are in what groups. We can do that by defining a design matrix. 
```{r}
design_matrix <- readr::read_csv("tests/testdata/design_matrix.csv")

head(design_matrix)
```
### Normalyzing data

mspms uses the NormalyzerDE package to do normalization under the hood.

We can normalyze the data as follows:  
```{r}
normalyzed_data <- normalyze(peaks_prepared_data, design_matrix)

```

### Handling Outliers

Here we use a Dixon test from the outliers package to detect outliers from each 
of our replicates. We need to know what samples are part of which groups, so we 
need to specify the design matrix here too. Make sure that the column header 
names are "sample" and "group" just like before.

```{r}
design_matrix <- readr::read_csv("tests/testdata/design_matrix.csv")
outliers <- handle_outliers(normalyzed_data, design_matrix)

```

### Imputation of data

We have a lot of missing, or 0 values. For these, we need to impute them so we 
can do downstream statistics.

```{r}
imputed <- mspms::impute(outliers)

```

### Joining with Library

Next we need to join everything with the sequences of the peptide library.  
```{r}
joined_with_library <- mspms::join_with_library(imputed)
```

### Calcuating cleavages.

Next, we need to determine where the peptide sequences are cleaved.

We check both the N and C terminus.

Sequences are presented as the user specified number of amino acids on 
both sides of a cleavage. The default is 4, but there is interest in looking at
motifs that are farther away from the cut site. Note that X indicates that there
was nothing at that position because it is over the end.

```{r}
cleavage_added_data <- mspms::add_cleavages(joined_with_library, n_residues = 4)

head(cleavage_added_data)
```

### Adding metadata back

We need to add information about our samples back to perform downstream 
statistics on the data. We accomplish this with the prepare for stats file.

```{r}
prepared_for_stats <- mspms::prepare_for_stats(cleavage_added_data,
                                               design_matrix)
```

### Polishing

This function polishes the data by combining the cterm and nterm cleavage 
information into one column while removing any rows that don't have any 
cleavage information or have cleavage information on the cterm and nterm.

```{r}
polished_data <- mspms::polish(cleavage_added_data)
```

### Complete workflow

There are a lot of steps to get to the final normalized data we listed above. 
We have combined all of these steps into one function, mspms() for convenience.
This offers a nice balance between allowing the user the option to use 
individual steps and providing a quick way to get to the final data.

```{r}
mspms_data <- mspms::mspms(peaks_prepared_data, design_matrix)
```

## Statistics

mspms provides a number of convenience functions to conduct statistics
on the data.

First, we need to prepare the data for stats. This involves reshaping the data 
into the long format so it is easy to conduct statistics on and then appending 
the data in the design matrix that allows us to determine what sample contains 
what conditions/time and conduct the appropriate statistics.

Now we can conduct the statistics:

### T tests

T tests are performed within each condition and compared to time 0.

For example, for an experiment where there are two conditions, 
DMSO and MZB as well as 3 time points, 0, 1, and 2, the t tests would be
as follows:

* DMSO.TO vs DMSO.T0. 
* DMSO.T0 vs DMSO.T1. 
* DMSO.T0 vs DMSO.T2.

* MZB.T0 vs MZB.T0. 
* MZB.T0 vs MZB.T1. 
* MZB.T0 vs MZB.T2.

```{r}
# Perform T test
t_test_stats <- mspms::mspms_t_tests(mspms_data)
```

### log2FC

We can also calculate the log 2 fold change. 
The comparisons here are the same as for the T tests.  
```{r}
# calculate log2fc
log2fc <- mspms::mspms_log2fc(mspms_data)

head(log2fc)
```

### log2fc_t_tests.

Oftentimes (such as when you want to make volcano plots) it is most useful to
look at the log2fc and the t test statistics at the same time.

```{r}
log2fc_t_test <- mspms::log2fc_t_test(mspms_data)

head(log2fc_t_test)
```

### log2fct_condition

You may wish to compare two conditions at a specific time point. 
This function allows this

```{r}
condition_t = mspms::log2fct_condition(mspms_data,
                                       ref_condition = "DMSO",
                                       comparison_condition = "MZB",
                                       at_time = 60)

head(condition_t)

```

### log2fct_time

You may wish to compare two time points within a specific condition. 
This function allows

```{r}
time_t = mspms::log2fct_time(mspms_data,
                             within_condition = "DMSO",
                             ref_time = 0,
                             comparison_time = 60)

head(time_t)
```

### ANOVA

We also might want to perform an anova. Here, we have it set up to show the 
effect of time within each condition.

For example, for an experiment where there are two conditions, DMSO and MZB as
well as 3 time points, 0, 1, and 2, the anova would statistics for the effect of
time for each peptide within DMSO or MZB.

```{r}
# Doing ANOVA
anova_stats <- mspms::mspms_anova(mspms_data)

head(anova_stats)
```

## Common Data Visualizations

We also provide some functions that make common data visualizations easier.

### volcano plots

```{r}
volcano_plot <- mspms::plot_volcano(log2fc_t_test)

volcano_plot
```

### Plotting sequence specificity motif

Here use an approach similar to what is implemented in ICELOGO to visualize the
sequence specificity of the cleavage sites. This was used as reference to build 
the code: https://iomics.ugent.be/icelogoserver/resources/manual.pdf.

Below we can see this applied to see what cleavage sequences are enriched among
the peptides significantly different between time 0 and time 60 
within DMSO.

```{r}
cleavage_seqs <- time_t %>%
  dplyr::filter(log2fc > 3, p.adj <= 0.05) %>%
  dplyr::pull(cleavage_seq)

background_universe <- mspms::all_possible_8mers_from_228_library

mspms::plot_icelogo(cleavage_seqs, background_universe)
```

We could also look at the fold change instead of the percent difference.
```{r}
mspms::plot_icelogo(cleavage_seqs, background_universe, type = "fold_change")
```

We also provide a convenience function for looking at all icelogos relative to 
time 0 within the experiment.

This function includes cleavages that were significant at any time point once in 
the experimental set of the ICELOGO. Below we show this applied to create
percent difference and fold change icelogos.

```{r}
plot_all_icelogos(mspms_data)

plot_all_icelogos(mspms_data, type = "fold_change")
```

### Extracting regular expressions from icelogos

It might be useful for the user to extract the information contained in the 
Icelogo and use it to filter their data for peptides that are predicted 
to be well cleaved, or not cleaved according to the icelogo plot.

For example, we can get the amino acids that are enriched at each position of 
the icelogo using the code below.

```{r}

# Getting data for Icelog 
cleavage_seqs <- time_t %>%
  dplyr::filter(log2fc > 3, p.adj <= 0.05) %>%
  dplyr::pull(cleavage_seq)

background_universe <- mspms::all_possible_8mers_from_228_library

# Plotting icelogo
mspms::plot_icelogo(cleavage_seqs, background_universe)

# corresponding icelogo data to the plot
icelogo_matrix = mspms::prepare_icelogo_data(cleavage_seqs, background_universe)

# Extracting regular expressions
RE_pos = mspms::extract_re(icelogo_matrix)

RE_pos
```

We can also do this for amino acids that are negatively enriched.

```{r}
RE_neg = mspms::extract_re(icelogo_matrix,type = "negative")

RE_neg
```

Once we have these regular expressions, we can use it to filter our data as
follows:

```{r}
pos_enriched = mspms_data %>% 
  dplyr::filter(grepl(RE_pos, cleavage_seq))
```

In this case, none of the peptides in the original data set match all the rules
in our library (peptide library does not contain all possible combinations, so
none of the peptides was a match to the combination of rules at each position)

We could then filter to only include peptides that are greater than a certain
threshold percent difference. We can then visualize these results easily with 
the plot_time_course function described further below. Here we will only 
consider AAs with a percent difference greater than 10 at each position.

```{r}
# Extracting regular expressions 
RE_pos = mspms::extract_re(icelogo_matrix,threshold = 10)

RE_pos

# Filtering data for peptides that have cleavage sequences matching regular 
# expressions
f = mspms_data %>% 
  dplyr::filter(grepl(RE_pos, cleavage_seq))

#visualizing these peptides that match over time.
plot_time_course(f)
```

Here we see that these peptides were cleaved over time in DMSO, as would be 
expected based on the icelogo plot.

### PCA.

We can generate a PCA plot to visualize the data. 
Here, the colors show the different time points while the shape shows 
the different conditions.

```{r}
mspms::plot_pca(mspms_data)
```

### Hierchical clustering

We can also generate an interactive heatmap with the hierarchical clustering 
results.

The format of this readme does not allow for interactivity, so a static picture 
of the output is shown instead.

```{r,eval = FALSE}
mspms::plot_heatmap(mspms_data)
```

![](README_files/plot_heatmap_output.png)

### Ploting time course.

We also provide a function for plotting the mean intensity and standard 
deviation over time for each peptide in the data set by conditions.
Facets show peptide, color shows the condition.

This can be used in combination with the ANOVA function.
Below, we will use anova to calculate the peptides where we see an 
effect of time (in the DMSO group), and then plot these.

```{r}

top10sig <- anova_stats %>%
  dplyr::arrange(p.adj) %>%
  dplyr::pull(Peptide) %>%
  head(10)

p1 <- mspms_data %>%
  dplyr::filter(Peptide %in% top10sig) %>%
  mspms::plot_time_course()

p1
```

## Calculating the number of cleavages at each position in the library

We can also calculate the number of cleavages at each position of the library.
Here we will do this for peptides significantly different relative to time 0 for
DMSO and MZB.

```{r}
# Performing T tests
t_tests = log2fc_t_test(mspms_data) %>% 
  dplyr::filter(p.adj < 0.05,
                log2fc > 3)

count = mspms::count_cleavages_per_pos(t_tests)

p1 = mspms:::plot_cleavages_per_pos(count)

p1
```