### 2024-09-19 README.Rmd
major structural changes after Bioconductor feedback

---
output: github_document
---

<!-- README.md is generated from README.Rmd. Please edit that file -->

```{r, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>"
)

ggplot2::theme_set(
  ggplot2::theme_minimal()
)

```

# mspms

<!-- badges: start -->

[![R-CMD-check](https://github.com/baynec2/mspms/actions/workflows/R-CMD-check.yaml/badge.svg)](https://github.com/baynec2/mspms/actions/workflows/R-CMD-check.yaml)
[![Codecov test
coverage](https://codecov.io/gh/baynec2/mspms/branch/main/graph/badge.svg)](https://app.codecov.io/gh/baynec2/mspms?branch=main)
[![BioC
status](http://www.bioconductor.org/shields/build/release/bioc/mspms.svg)](https://bioconductor.org/checkResults/release/bioc-LATEST/mspms)

<!-- badges: end -->

The goal of mspms is provide a R Package that can be used to analyze data from
the [Multiplex Substrate Profiling by Mass Spectrometry (MSP-MS)
method](https://pubmed.ncbi.nlm.nih.gov/36948708/).

Additionally, we provide a [graphical user interface powered by shiny
apps](https://gonzalezlab.shinyapps.io/mspms_shiny/) that allows for a user to
utilize the method without requiring any R coding knowledge.

## Installation

You can install the development version of mspms from github.

```{r,eval = FALSE}
devtools::install_github("baynec2/mspms",
                         ref = "develop")
```

## Quickstart

To generate a general report using your own data, run the following code. It
requires data that has been prepared for mspms data analysis by a converter
function, and a design matrix. For more information see subsequent sections.

```{r,eval = FALSE}
mspms::generate_report(
  prepared_data = mspms::peaks_prepared_data,
  colData = mspms::design_matrix,
  outdir = "../Desktop/mspms_report"
)
```

The above command will generate a .html file containing a generic mspms
analysis.

There is much more that can be done using the mspms package- see the following
sections for more information.

## Overview

Functions in this package are logically divided into 4 broad types of functions.

1.  Pre-processing data. These functions are focused on getting all of the data
    necessary for a mspms analysis together and in a consistent format.

2.  Data processing. These functions allow the user to process the proteomics
    data (imputation, normalization, etc).

3.  Statistics. These methods allow the user to perform basic statistics on the
    normalized/processed data.

4.  Data visualization. These functions allow the user to visualize the data.

**Note that only a subset of functions are exported to the user.** Helper
functions can be found in the \_helper_functions.R files.

## Pre-processing data

Pre-processing data here refers to standardizing the data exported from
different software solutions for analyzing mass spectrometry-based proteomics
and adding useful information specific to the MSP-MS assay.

Specifically, the data from the file exported from the upstream software is
parsed to contain the **peptide** (detected peptide sequence, cleavages are
represented as "\_"), **library_id** (name of the peptide in the peptide library
each detected peptide maps to), and intensities for each sample.

Information about each peptide relative to the peptide it was derived from is
then added. This includes the **cleavage_seq** (the motif the peptide a user
specified n residues up and down from the cleavage event), **cleavage_pos**
(position of the peptide library the cleavage product was cleaved at ),
**peptide_type** (whether the detected peptide is a cleavage product of the
peptide library, or a full length member of the peptide library).

All of the information is then loaded as a QFeatures object along with colData
(metadata about each sample) where it is stored as a SummarizedExperiment object
named "peptides".

#### Uploading your colData

As briefly described above, the colData contains metadata describing the samples
you are trying to analyze.

It is very important that this contains at least the following columns
-"quantCols","group","condition",and "time".

The quantCols field contains the name of each sample. These names must match
whatever the samples are named in the files exported from the upstream
proteomics software.

The group field contains text specifying what group each sample belongs to. All
replicate samples should have the same group name

The condition field contains the experimental conditions that vary in your
MSP-MS experiment (this depends on the experimental question, could be type of
protease, the type of inhibitor, etc)

The time field contains the duration of incubation of your protease of interest
with the peptide library for each sample.

An example of a valid colData file is shown below.

```{r}
colData = readr::read_csv("inst/extdata/colData.csv")
usethis::use_data(colData,overwrite = TRUE)

head(colData)
```

#### PEAKS

Analysis using a .csv exported from [PEAKS
software](https://www.bioinfor.com/peaks-studio/) is supported.

an exported .csv file can be generated from PEAKS as follows:

1.  Create a New Project in PEAKS.\
2.  Select data to add to PEAKS project.\
3.  Add data, rename samples, set correct parameters. • Highlight all the added
    data.\
    • Select Create new sample for each file.\
    • Select appropriate instrument and fragmentation method.\
    • Set Enzyme to “None”.\
    • Rename samples according to enzymes, time points, and replicates.\
    • Can also ensure appropriate data is selected for these.\
    • Select “Data Refinement” to continue.\
4.  Data Refinement.\
    • On the top right, select MSP-MS as the predefined parameters.\
    • Parameters shown are what should be used for MSP-MS data analysis.\
    • Select “Identification” to continue.\
5.  Peptide Identification.\
    • On the top right, select MSP-MS as the predefined parameters.\
    • Parameters shown are what should be used for MSP-MS data analysis.\
    • Identification -\> Select Database -\> TPD_237.\
    • Identification -\> no PTMs.\
    • Can remove the PTMs by highlighting them and selecting Remove\
6.  Label Free Quantification.\
    • Make sure Label Free is selected.\
    • Group samples -\> add quadruplicates of samples to new group.

```{r}
library(dplyr)
library(mspms)

# File path of peaks lfq file
lfq_filename <- "inst/extdata/peaks_protein-peptides-lfq.csv"

# Prepare the data 
peaks_prepared_data <- mspms::prepare_peaks(lfq_filename,
                                            colData,
                                            0.3,
                                            n_residues = 4)

usethis::use_data(peaks_prepared_data,overwrite = TRUE)

peaks_prepared_data
```

#### Proteome Discoverer

Analysis using a .csv file exported from [proteome
discoverer](https://www.thermofisher.com/us/en/home/industrial/mass-spectrometry/liquid-chromatography-mass-spectrometry-lc-ms/lc-ms-software/multi-omics-data-analysis/proteome-discoverer-software.html)
is supported.

To generate this .csv file:

```{r}

```

#### Fragpipe

Data exported from [Fragpipe](https://github.com/Nesvilab/FragPipe) as a .tsv
file is supported.

To generate this file, run a [LFQ Fragpipe
workflow](https://fragpipe.nesvilab.org/docs/tutorial_lfq.html) using your
desired parameters (we recommend using match between runs).

You will find a file named "combined_peptide.tsv" in the output folder from
Fragpipe.

This can be loaded into mspms as follows:

```{r}
lfq_file_path = "inst/extdata/fragpipe_combined_peptide.tsv"

fragpipe_prepared_data = prepare_fragpipe(lfq_file_path,
                                          colData)

fragpipe_prepared_data
```

## Data Processing

Data processing includes imputation and normalization of the proteomics data.

**Add more info**

```{r}
# does set seed matter?
set.seed(2)
processed_qf = process_qf(peaks_prepared_data)

usethis::use_data(processed_qf,overwrite = TRUE)

processed_qf
```

## Statistics

MSP-MS analysis has traditionally used T-tests to what peptides are
significantly different relative to T0. The authors of the assay suggest using a
T-test derived FDR adjusted p value threshold \<= 0.05 and log2 fold change
threshold \>= 3.

We provide a user-friendly implementation of this statistical test as follows:

```{r}
log2fc_t_test_data = mspms::log2fc_t_test(processed_qf = processed_qf)

usethis::use_data(log2fc_t_test_data,overwrite = TRUE)
```

Please note that statistical analyses are one of the strengths of the
Bioconductor project and several packages provide alternative statistical
approaches that are more flexible than t-tests and could be used with the
QFeatures objects produced by mspms.

This includes packages using empirical Bayesian approaches (see
[limma](https://bioconductor.org/packages/release/bioc/html/limma.html)) and
mixed modelling (see
[msqrob2](https://bioconductor.org/packages/release/bioc/html/msqrob2.html)),
that are suited for proteomics data.

Users are encouraged to use these excellent packages.

## Data Visualizations

We provide a number of data visualizations as part of the mspms_package.

### QC Checks

A critical initial step in MSP-MS data analysis is a rigorous quality control
assessment. Given the reliance on a specific peptide library, calculating the
percentage of library peptides detected in each sample provides a
straightforward yet powerful indicator of data integrity. Anomalously low
detection rates signal potential instrument or experimental problems.

We can see the percentage of the peptide library that were undetected among each
group of replicates as follows (for full length peptides a threshold of \~\< 10
is considered good while for cleavage products a threshold \~\< 5 is good):

```{r}
plot_qc_check(processed_qf,
              full_length_threshold = 10,
              cleavage_product_threshold = 5)
```

We can also see in what percentage of samples each peptide in the peptide
library were undetected in.

```{r}
plot_nd_peptides(processed_qf)
```

### Heatmap

We can inspect our data using an interactive heatmap as follows:

```{r}
plot_heatmap(processed_qf)
```

### PCA

We can also inspect our data using a PCA as follows.

```{r}
plot_pca(processed_qf)
```

### Volcano plots

We can generate volcano plots showing the significant peptides from each
condition as follows:

```{r}
plot_volcano(log2fc_t_test_data)
```

### Cleavage Position Plots

We can also create plots showing the number of times we observe a cleavage at
each position of the peptide_library among a set of peptides of interest.

We can do this as follows:

```{r}
sig_cleavage_data = log2fc_t_test_data %>% 
  dplyr::filter(p.adj <= 0.05,log2fc > 3)

p1 = mspms::plot_cleavages_per_pos(sig_cleavage_data )

p1
```

### iceLogo

[Icelogos](https://www.nature.com/articles/nmeth1109-786) are a method for
visualizing conserved patterns in protein sequences through probability theory.

As part of the mspms R package, we have implemented the logic to generate
[iceLogos](https://genesis.ugent.be/uvpublicdata/icelogo/iceLogo.pdf) in R.

To use this, we first have to define a vector of cleavage sequences that we are
interested in. We will do this for the significant CatA cleavages below.

```{r}
catA_sig_cleavages = log2fc_t_test_data %>% 
  dplyr::filter(p.adj <= 0.05,log2fc > 3) %>% 
  dplyr::filter(condition == "CatA") %>% 
  dplyr::pull(cleavage_seq) %>% 
  unique()
```

We also need to know all of the possible cleavages we could see among our
peptide library. We can generate that as below:

```{r}
all_possible_8mers_from_228_library = calculate_all_cleavages(mspms::peptide_library$library_real_sequence,n_AA_after_cleavage = 4)

usethis::use_data(all_possible_8mers_from_228_library,overwrite = TRUE)
```

We can then generate an iceLogo relative to the background of all possible
cleavages in our peptide library like so:

```{r}
plot_icelogo(catA_sig_cleavages,
             background_universe = all_possible_8mers_from_228_library)
```

We also provide a function that conveniently plots all icelogos from all
conditions from the experiment all together.

```{r,fig.height = 10, fig.width = 4}
sig_cleavage_data = log2fc_t_test_data %>% 
  dplyr::filter(p.adj <= 0.05,log2fc > 3)

plot_all_icelogos(sig_cleavage_data)
```

### Plotting a Time Course

We can also plot the intensities of peptides of interest over time.

First we need to define which peptides we are interested in. Here we will use
the t-test results, and look at the peptide with the max log2fc.

```{r}
max_log2fc_pep = log2fc_t_test_data %>% 
  dplyr::filter(p.adj <= 0.05,log2fc > 3) %>% 
  dplyr::filter(log2fc ==max(log2fc)) %>% 
  pull(peptide)

```

We can then filter our QFeatures object to only include this peptide

```{r}
filter = QFeatures::VariableFilter(field = "peptide",
               value = max_log2fc_pep,
               condition = "==")

sel = QFeatures::filterFeatures(processed_qf,filter)
```

And then plot as follows:

```{r}
plot_time_course(sel,summarize_by_mean = FALSE)
```